---
title: "Practical Machine Learning - Assignment"
author: "Koushik Nagarajan"
date: "22 June 2015"
output: html_document
---

##Summary
Creating a model or finding a pattern from a large dataset from a group of enthusiasts who take measurements about themselves regularly to improve their health. This model can be used to predict any new scenario and find out the pattern (Classe)

##Data Loading

```{r cache=TRUE}
library(caret)
trainingData <- read.csv("data/pml-training.csv",na.strings = c("NA", ""))
testingData <- read.csv("data/pml-testing.csv", na.strings = c("NA", ""))
 
```

##Data Cleansing

###Near Zero Variance
```{r}
myDataNZV <- nearZeroVar(trainingData)
trainingData <- trainingData[,-myDataNZV]
testingData <- testingData[,-myDataNZV]
```
###Missing Columns
```{r}
nacolumns <- (colSums(is.na(trainingData)) == 0)
trainingData <- trainingData[, nacolumns]
testingData <- testingData[, nacolumns]
```
###Ignored Columns
```{r}
otherColumnsToIgnore <- grepl("X|user_name|timestamp|new_window", colnames(trainingData))
trainingData <- trainingData[, !otherColumnsToIgnore]
testingData <- testingData[, !otherColumnsToIgnore]
```

##Data Partition
```{r}
inTrain <- createDataPartition(y=trainingData$classe, p=0.6, list=FALSE)
myTraining <- trainingData[inTrain, ]; myTesting <- trainingData[-inTrain, ]

```
##Data Analysis
```{r}
dim(myTraining); dim(myTesting)
```
##Building the Model

###Chosing the model
```{r}
library(randomForest)
modFit <- randomForest(classe ~. , data=myTraining)
predictions <- predict(modFit, myTesting, type = "class")
```
###Cross Validation Testing and Out-of-Sample Error Estimate
```{r}
confusionMatrix(predictions, myTesting$classe)
```
##Predicted Results
```{r}
predictions <- predict(modFit, testingData, type = "class")

pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(predictions)

```